---
title: "Document_Outputs"
author: "Nick Zhang"
format: pdf
editor: visual
---

## COM-Poisson football model

Markdown to obtain all the relevant plots and tables of the report on

#### Imports

```{r}
source('utils.R') # Contains all utility functions to support modelling and analysis, including functions to import/export data, manipulate the MH_posterior objects from the MCMC routines, and generate predictions
source('MH_CMPSAS_STZ.R')


library('ggplot2')
library('tidyverse')

library('patchwork')
library('matrixStats')
```

#### SAS Prior Density

```{r}
w_spike <- 0.3
w_slab  <- 0.7
sigma_spike <- 0.05
sigma_slab  <- 1
x <- seq(-2, 2, length.out = 5000)  # more points = smoother line
mixture <- w_spike * dnorm(x, 0, sigma_spike) + w_slab * dnorm(x, 0, sigma_slab)

# Data for both scales
df_log <- data.frame(x = x, mixture = mixture)
df_exp <- data.frame(nu = exp(x), mixture = mixture)

# Custom sand-like color
sand_color <- "#c2a15e"

# --- Plot 1: Spike and slab on log(ν) scale ---
p1 <- ggplot(df_log, aes(x, y = mixture)) +
  geom_line(color = sand_color, size = 0.8, lineend = "round") +
  labs(
    title = "Spike and Slab Prior Density (log scale)",
    x = expression(log(nu)),
    y = "Density"
  ) +
  theme_minimal(base_size = 10)

# --- Plot 2: Same density on ν = exp(log(ν)) scale ---
p2 <- ggplot(df_exp, aes(x = nu, y = mixture)) +
  geom_line(color = sand_color, size = 0.8, lineend = "round") +
  coord_cartesian(xlim = c(0, 7)) +
  labs(
    title = "Spike and Slab Prior Density (original scale)",
    x = expression(nu),
    y = "Density"
  ) +
  theme_minimal(base_size = 10)

combined_plot = p1 / p2

```

#### Exploratory Data Analysis

```{r}

X = read_data('2324', 'Premier')
X1 = X[[1]]
X2 = X[[2]]

# Remove diagonals
X1_nodiag <- X1[upper.tri(X1) | lower.tri(X1)]
X2_nodiag <- X2[upper.tri(X2) | lower.tri(X2)]

# Truncate values > 7
X1_nodiag <- X1_nodiag[X1_nodiag <= 7]
X2_nodiag <- X2_nodiag[X2_nodiag <= 7]

# Convert to factors for exact bins 0 to 7
X1_factor <- factor(X1_nodiag, levels = 0:7)
X2_factor <- factor(X2_nodiag, levels = 0:7)

# Calculate means for vertical lines
mean_X1 <- mean(as.numeric(as.character(X1_factor)))
mean_X2 <- mean(as.numeric(as.character(X2_factor)))
var_X1 <- var(as.numeric(as.character(X1_factor)))
var_X2 <- var(as.numeric(as.character(X2_factor)))

# Create individual plots

p1 <- ggplot(data.frame(value = X1_factor), aes(x = value)) +
  geom_bar(fill = "#87CEEB", color = "black", width = 1) +
  geom_vline(xintercept = mean_X1 + 1, linetype = "solid", color = "#0072B2", linewidth = 1) +
  geom_vline(xintercept = var_X1 + 1, linetype = "dashed", color = "#8E44AD", linewidth = 0.8) +
  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Home Goals Scored", y = "Number of Matches") +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 10),
    plot.margin = margin(t = 5, r = 10, b = 5, l = 5)
  )

p2 <- ggplot(data.frame(value = X2_factor), aes(x = value)) +
  geom_bar(fill = "#87CEEB", color = "black", width = 1) +
  geom_vline(xintercept = mean_X2 + 1, linetype = "solid", color = "#0072B2", linewidth = 1) +
  geom_vline(xintercept = var_X2 + 1, linetype = "dashed", color = "#8E44AD", linewidth = 0.8) +
  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Away Goals Scored", y = NULL) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    plot.margin = margin(t = 5, r = 5, b = 5, l = 10)
  )

# Get max y-axis limit to sync y scales
max_y <- max(ggplot_build(p1)$data[[1]]$count, ggplot_build(p2)$data[[1]]$count)

# Fix y-axis limits to be the same
p1 <- p1 + coord_cartesian(ylim = c(0, max_y))
p2 <- p2 + coord_cartesian(ylim = c(0, max_y))

# Combine with patchwork, add space between plots
combined_plot <- p1 + p2 + plot_layout(ncol = 2, widths = c(1, 1)) & 
  theme(plot.margin = margin(10, 10, 10, 10)) 

print(combined_plot)
# H:800 W:350
print(mean(X1_nodiag))
print(mean(X2_nodiag))
print(var(X1_nodiag))
print(var(X2_nodiag))
```

```{r}

```

```{r}
season = '2324'
league_acro = 'PL'
league = 'Premier'

# X = read_data('2122', 'WSL')

X = read_data(season, league)
X1 = X[[1]]
X2 = X[[2]]

df_hist = read_historics(season, league_acro)
team_names = sort(unique(df_hist[,"HomeTeam"]))
N = length(team_names)
rownames(X1) = team_names
rownames(X2) = team_names
team_codes = setNames(1:N, team_names)

for (j in 1:N){
  goals = unname(append(X1[j,],X2[,j]))
  goals = goals[!is.na(goals)]
  cat(team_names[j],mean(goals)/var(goals),'\n')
}

plots = list()
for (j in 1:N){
  # goals = unname(append(X1[j,],X2[,j]))
  goals = unname(append(X1[j,],X2[,j]))
  goals = goals[!is.na(goals)]
  
  if (mean(goals)/var(goals) > 1){disp_color = '#D55E00'}else{disp_color = '#009E73'}
  
  mean_goals = mean(goals)
  var_goals = var(goals)
  
  p = ggplot(data.frame(goals), aes(x = goals)) +
    coord_cartesian(xlim=c(0,7), ylim=c(0,15)) +
    geom_histogram(binwidth = 1, fill = disp_color, color = 'black') +
    geom_vline(xintercept = mean_goals, linetype = "solid", color = "#0072B2", linewidth = 0.8) +
    geom_vline(xintercept = var_goals, linetype = "dashed", color = "#8E44AD", linewidth = 0.6) +
    # geom_vline(xintercept = mean_goals, color = '#0072B2', linetype = 'solid', linewidth =0.8, alpha = 1) +
    theme_minimal() +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
    # labs(x = 'Goals', y ='Number of Matches') +
    ggtitle(paste(team_names[j]))
  
  plots[[j]] = p
}

# do.call("grid.arrange", c(plots[1:10], nrow = 5, ncol = 2, left = 'counts', bottom = 'goals'))
# do.call("grid.arrange", c(plots[11:20], nrow = 5, ncol = 2, left = 'counts', bottom = 'goals'))

# H:900 W:550
do.call("grid.arrange", c(plots, nrow = 4, ncol = 5, left = 'Number of Matches', bottom = 'Goals'))
```

##### Appendix: Defense plots

```{r}

# seasons_strvec = generate_season_string(2014, 2025)
# tot_h = 0
# tot_a = 0
# for (season in seasons_strvec){
#   print(season)
#   league_acro = 'LL'
#   league = 'Liga'
#   
#   
#   X = read_data(season, league)
#   X1 = X[[1]]
#   X2 = X[[2]]
#   
#   df_hist = read_historics(season, league_acro)
#   team_names = sort(unique(df_hist[,"HomeTeam"]))
#   rownames(X1) = team_names
#   rownames(X2) = team_names
#   team_codes = setNames(1:20, team_names)
#   h_c = 0
#   a_c = 0
#   for (j in 1:20){
#     goals = unname(append(X1[j,],X2[,j]))
#     goals = goals[!is.na(goals)]
#     h_c = h_c + abs(1-mean(goals)/var(goals))
#     # cat(team_names[j],abs(1-mean(goals)/var(goals)),'\n')
#     # 
#     goals = unname(append(X1[,j],X2[j,]))
#     goals = goals[!is.na(goals)]
#     # cat(team_names[j],abs(1-mean(goals)/var(goals)),'\n')
#     a_c = a_c + abs(1-mean(goals)/var(goals))
#   }
#   tot_h = tot_h + h_c/20
#   tot_a = tot_a + a_c/20
#   print(h_c / 20)
#   print(a_c / 20)
# }
# print(tot_h / length(seasons_strvec))
# print(tot_a / length(seasons_strvec))

season = '2324'
league_acro = 'PL'
league = 'Premier'


X = read_data(season, league)
X1 = X[[1]]
X2 = X[[2]]

df_hist = read_historics(season, league_acro)
team_names = sort(unique(df_hist[,"HomeTeam"]))
rownames(X1) = team_names
rownames(X2) = team_names
team_codes = setNames(1:20, team_names)
plots = list()
for (j in 1:20){
  # goals = unname(append(X1[j,],X2[,j]))
  goals = unname(append(X1[,j],X2[j,]))
  goals = goals[!is.na(goals)]
  
  if (mean(goals)/var(goals) > 1){disp_color = '#D55E00'}
  else{disp_color = '#009E73'}
  
  mean_goals = mean(goals)
  var_goals = var(goals)
  
  p = ggplot(data.frame(goals), aes(x = goals)) +
    coord_cartesian(xlim=c(0,7), ylim=c(0,15)) +
    geom_histogram(binwidth = 1, fill = disp_color, color = 'black') +
    geom_vline(xintercept = mean_goals, linetype = "solid", color = "#0072B2", linewidth = 0.8) +
    geom_vline(xintercept = var_goals, linetype = "dashed", color = "#8E44AD", linewidth = 0.6) +
    # geom_vline(xintercept = mean_goals, color = '#0072B2', linetype = 'solid', linewidth =0.8, alpha = 1) +
    theme_minimal() +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
    ggtitle(paste(team_names[j]))
  
  plots[[j]] = p
}

# do.call("grid.arrange", c(plots[1:10], nrow = 5, ncol = 2, left = 'counts', bottom = 'goals'))
# do.call("grid.arrange", c(plots[11:20], nrow = 5, ncol = 2, left = 'counts', bottom = 'goals'))

do.call("grid.arrange", c(plots, nrow = 4, ncol = 5, left = 'Number of Matches', bottom = 'Goals'))
```

```{r}
plots <- list()

for (j in 1:20) {
  # Combine row and column for team j
  goals <- unname(c(X1[j, ], X2[, j]))
  goals <- goals[!is.na(goals)]
  
  disp_color <- if (mean(goals) / var(goals) > 1) "#D55E00" else "#009E73"
  mean_goals <- mean(as.numeric(as.character(goals))) 
  # Truncate to 0–7
  goals[goals > 6] <- 6
  
  # Convert to factor with levels 0:7 to ensure all bins shown
  goals_factor <- factor(goals, levels = 0:6)
  
  # Create plot
  p <- ggplot(data.frame(goals = goals_factor), aes(x = goals)) +
    geom_bar(fill = disp_color, color = 'black', width = 1) +
    geom_vline(xintercept = mean_goals + 1,  # +1 since factor levels are centered at 1-based index
               color = '#0072B2', linetype = 'solid', linewidth = 0.8, alpha = 1) +
    scale_x_discrete(drop = FALSE, labels = 0:6) +
    coord_cartesian(ylim = c(0, 15)) +
    theme_minimal() +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text = element_text(size = 8),
      plot.title = element_text(size = 10, hjust = 0.5)
    ) +
    ggtitle(team_names[j])
  
  plots[[j]] <- p
}

# Combine plots in a grid
do.call("grid.arrange", c(plots, nrow = 4, ncol = 5,
                          left = 'Counts', bottom = 'Goals'))
```

#### Posterior Analysis

##### Att and Mean scatter plot comparison

```{r}
league_acro = 'PL'
season = '2324'
n_run = 1

library(matrixStats)
load(file = paste0("Data//MH_Results//Pois_FullLeague//",league_acro,"_",season,"_Run",n_run,"_Pois",".RData"))

# MH_object_pois = MH_P
# MH_object_pois = thin_mcmc(MH_P, 4)
t = MH_object_pois$iterations
df_plot = data.frame(att_means = colMedians(MH_object_pois$att_post[(t/5):t,]),
                     def_means = colMedians(MH_object_pois$def_post[(t/5):t,]),
                     label = MH_object_pois$team_names)

# load(file = paste0("Data//MH_Results//SAS_FullLeague//",league_acro,"_",season,"_Run",n_run,"_SAS",".RData"))
load(file = paste0("Data//MH_Results//SAS_FullLeague//",league_acro,"_",season,"_Run1_SAS",".RData"))

MH_object = MH_SAS
# MH_object = thin_mcmc(MH_SAS, 10)
t = MH_object$iterations
df_plot2 = data.frame(att_means = colMedians(MH_object$att_post[(t/5):t,]),
                      def_means = colMedians(MH_object$def_post[(t/5):t,]),
                      label = MH_object$team_names)

team_names = MH_object$team_names
```

```{r}
df_colors <- data.frame(
  name = team_names,
  color = c(
    "#00539F", # Atalanta
    "#A5051A", # Bologna
    "#8B0000", # Cagliari
    "#0072CE", # Como
    "#2A52BE", # Empoli
    "#5F259F", # Fiorentina
    "#D20000", # Genoa
    "#1E90FF", # Inter
    "#000000", # Juventus
    "#A2D2FF", # Lazio
    "#FFD700", # Lecce
    "#DA291C", # Milan
    "#E10600", # Monza
    "#0076BE", # Napoli
    "#FFD700", # Parma
    "#8E1B1B", # Roma
    "#781F1F", # Torino
    "#2E2E2E", # Udinese
    "#FF6A13", # Venezia
    "#FFD100"  # Verona
  )
)

df_plot <- cbind(df_plot, colors = df_colors$color)
df_plot2 <- cbind(df_plot2, colors = df_colors$color)
```

```{r}
df_colors <- data.frame(
  name = team_names,
  color = c(
    "#EF0107",  # Arsenal – classic red :contentReference[oaicite:1]{index=1}
    "#95BFE5",  # Aston Villa – claret alternative; not confirmed in sources
    "#DA291C",  # Bournemouth – cherry red (approx.)
    "#E7000F",  # Brentford – red (approx.)
    "#0057B8",  # Brighton – blue (approx.)
    "#034694",  # Chelsea – royal blue (common hex)
    "#1B458F",  # Crystal Palace – blue (approx.)
    "#003399",  # Everton – royal blue (approx.)
    "#000000",  # Fulham – black
    "#85BFEA",  # Ipswich – sky blue (approx.)
    "#003090",  # Leicester – blue (approx.)
    "#C8102E",  # Liverpool – red (approx.)
    "#6CABDD",  # Man City – sky blue (approx.)
    "#DA291C",  # Manchester United – red (approx.)
    "#241F20",  # Newcastle – black (approx.)
    "#E53236",  # Nottingham Forest – red (approx.)
    "#D71920",  # Southampton – red (approx.)
    "#132257",  # Tottenham – navy blue (approx.)
    "#7A263A",  # West Ham – claret (approx.)
    "#F5AF00"   # Wolves – gold (approx.)
  )
)
df_colors <- data.frame(
  name = team_names,
  color = c(
    "#EF0107",  # Arsenal – red
    "#95BFE5",  # Aston Villa – light blue alt
    "#DA291C",  # Bournemouth – red
    "#E7000F",  # Brentford – red
    "#0057B8",  # Brighton – blue
    "#6C1D45",  # Burnley – claret
    "#034694",  # Chelsea – royal blue
    "#1B458F",  # Crystal Palace – blue
    "#003399",  # Everton – royal blue
    "#000000",  # Fulham – black
    "#C8102E",  # Liverpool – red
    "#FF7F00",  # Luton – orange
    "#6CABDD",  # Man City – sky blue
    "#DA291C",  # Man United – red
    "#241F20",  # Newcastle – black
    "#E53236",  # Nott'm Forest – red
    "#EE2737",  # Sheffield United – red
    "#132257",  # Tottenham – navy blue
    "#7A263A",  # West Ham – claret
    "#F5AF00"   # Wolves – gold
  )
)
df_plot <- cbind(df_plot, colors = df_colors$color)
df_plot2 <- cbind(df_plot2, colors = df_colors$color)
```

```{r}
# df_colors <- data.frame(
#   name = team_names,
#   color = c(
#     "#000000",  # Arsenal – classic red :contentReference[oaicite:1]{index=1}
#     "#95BFE5",  # Aston Villa – claret alternative; not confirmed in sources
#     "#DA291C",  # Bournemouth – cherry red (approx.)
#     "#E7000F",  # Brentford – red (approx.)
#     "#0057B8",  # Brighton – blue (approx.)
#     "#034694",  # Chelsea – royal blue (common hex)
#     "#1B458F",  # Crystal Palace – blue (approx.)
#     "#003399",  # Everton – royal blue (approx.)
#     "#000000",  # Fulham – black
#     "#85BFEA",  # Ipswich – sky blue (approx.)
#     "#003090",  # Leicester – blue (approx.)
#     "#C8102E",  # Liverpool – red (approx.)
#     "#6CABDD",  # Man City – sky blue (approx.)
#     "#DA291C",  # Manchester United – red (approx.)
#     "#241F20",  # Newcastle – black (approx.)
#     "#E53236",  # Nottingham Forest – red (approx.)
#     "#D71920",  # Southampton – red (approx.)
#     "#132257",  # Tottenham – navy blue (approx.)
#     "#7A263A",  # West Ham – claret (approx.)
#     "#F5AF00"   # Wolves – gold (approx.)
#   )
# )

df_colors <- data.frame(
  name = team_names,
  color = c(
    "gray",  # Arsenal – classic red :contentReference[oaicite:1]{index=1}
    "gray",  # Aston Villa – claret alternative; not confirmed in sources
    "#009E73",  # Bournemouth – cherry red (approx.)
    "gray",  # Brentford – red (approx.)
    "gray",  # Brighton – blue (approx.)
    "gray",  # Chelsea – royal blue (common hex)
    "gray",  # Crystal Palace – blue (approx.)
    "gray",  # Everton – royal blue (approx.)
    "gray",  # Fulham – black
    "gray",  # Ipswich – sky blue (approx.)
    "gray",  # Leicester – blue (approx.)
    "#D55E00",  # Liverpool – red (approx.)
    "gray",  # Man City – sky blue (approx.)
    "#009E73",  # Manchester United – red (approx.)
    "gray",  # Newcastle – black (approx.)
    "gray",  # Nottingham Forest – red (approx.)
    "#D55E00",  # Southampton – red (approx.)
    "#009E73",  # Tottenham – navy blue (approx.)
    "#D55E00",  # West Ham – claret (approx.)
    "#D55E00"   # Wolves – gold (approx.)
  )
)

df_colors <- data.frame(
  name = team_names,
  color = c(
    "gray",  # Arsenal – classic red :contentReference[oaicite:1]{index=1}
    "gray",  # Aston Villa – claret alternative; not confirmed in sources
    "#009E73",  # Bournemouth – cherry red (approx.)
    "gray",  # Brentford – red (approx.)
    "#009E73",  # Brighton – blue (approx.)
    "gray",  # Chelsea – royal blue (common hex)
    "#009E73",  # Crystal Palace – blue (approx.)
    "gray",  # Everton – royal blue (approx.)
    "gray",  # Fulham – black
    "gray",  # Ipswich – sky blue (approx.)
    "#009E73",  # Leicester – blue (approx.)
    "#009E73",  # Liverpool – red (approx.)
    "gray",  # Man City – sky blue (approx.)
    "gray",  # Manchester United – red (approx.)
    "#009E73",  # Newcastle – black (approx.)
    "#D55E00",  # Nottingham Forest – red (approx.)
    "red3",  # Southampton – red (approx.)
    "#D55E00",  # Tottenham – navy blue (approx.)
    "gray",  # West Ham – claret (approx.)
    "gray"   # Wolves – gold (approx.)
  )
)

df_colors <- data.frame(
  name = team_names,
  color = c(
    "gray",  # Arsenal – classic red :contentReference[oaicite:1]{index=1}
    "gray",
    "gray",
    "gray",
    "#009E73",# Aston Villa – claret alternative; not confirmed in sources
    "gray",  # Brentford – red (approx.)
    "gray",  # Brighton – blue (approx.)
    "gray",  # Chelsea – royal blue (common hex)
    "gray",
    "#009E73",
    "gray",# Crystal Palace – blue (approx.)
    "gray",
    "gray",# Everton – royal blue (approx.)
    "gray",  # Fulham – black
    "gray",
    "#D55E00",# Ipswich – sky blue (approx.)
    "gray",  # Leicester – blue (approx.)
    "gray",  # Man City – sky blue (approx.)
    "gray",  # Newcastle – black (approx.)
    "#009E73"  # Nottingham Forest – red (approx.)
  )
)
df_plot <- cbind(df_plot, colors = df_colors$color)
# df_plot <- cbind(df_plot, colors = rep("gray", 20))
df_plot2 <- cbind(df_plot2, colors = df_colors$color)
```

```{r}
library(ggrepel)
p1 =  ggplot(df_plot, aes(x = att_means, y = def_means, label = team_names)) +
  geom_jitter(aes(color = colors), shape = 19, size = 2) + geom_text_repel(size = 3, max.overlaps = 10) +
  scale_color_identity() +
  scale_y_reverse() +
  # theme(axis.text.x = element_text(size = 10), axis.title.x = element_text(size = 14, face = 'bold')) +
  theme_minimal() + 
  geom_vline(xintercept = 0, linetype = 'dashed', alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = 'dashed', alpha = 0.2) +
  coord_cartesian(xlim = c(-1.5, 1)) +
  scale_y_continuous(limits = c(-1, 1)) + 
  ylab('') + xlab('') +
  theme(plot.title = element_text(face = "bold")) +
  ggtitle('Poisson Model')
p2 = ggplot(df_plot2, aes(x = att_means, y = def_means, label = team_names)) +
  geom_jitter(aes(color = colors), shape = 19, size = 2) + geom_text_repel(size = 3, max.overlaps = 10) +
  scale_color_identity() +
  scale_y_reverse() +
  # theme(axis.text.x = element_text(size = 10), axis.title.x = element_text(size = 14, face = 'bold')) +
  theme_minimal() + 
  geom_vline(xintercept = 0, linetype = 'dashed', alpha = 0.2) +
  geom_hline(yintercept = 0, linetype = 'dashed', alpha = 0.2) +
  coord_cartesian(xlim = c(-1.5, 1)) +
  scale_y_continuous(limits = c(-1, 1)) +
  ylab('') + xlab('') +
  theme(plot.title = element_text(face = "bold")) +
  ggtitle('CMP-SAS Model')

plots = list()
plots[[1]] = p1
plots[[2]] = p2


# do.call("grid.arrange", c(plots, nrow = 1, ncol = 2, left = 'Mean DEF', bottom = 'Mean ATT'))
```

```{r}
do.call("grid.arrange", c(plots, nrow = 1, ncol = 2, left = 'Mean DEF', bottom = 'Mean ATT'))
```

```{r}
league_acro = 'SA'
season = '2425'
n_run = 1
load(file = paste0("Data//MH_Results//Pois_FullLeague//",league_acro,"_",season,"_Run",n_run,"_Pois",".RData"))

MH_object_pois = MH_P
t = MH_object_pois$iterations
df_plot = data.frame(att_means = colMeans(MH_object_pois$att_post[(t/2):t,]),
                     def_means = colMeans(MH_object_pois$def_post[(t/2):t,]),
                     label = MH_object$team_names)
# mean(MH_object_pois$home_post[(t/2):t])
```

###### Home Coefficient

```{r}
league_acro = 'PL'

seasons_strvec = generate_season_string(2020, 2025)
plot_df = data.frame(matrix(ncol = 5, nrow = 0))
for (i in 1:length(seasons_strvec)){
  season = seasons_strvec[i]
  # team_fix_id = seas_fix[[i]][[2]]
  load(file = paste0("Data//MH_Results//SAS_FullLeague//",league_acro,"_",season,"_Run",n_run,"_SAS",".RData"))
  load(file = paste0("Data//MH_Results//CMP_FullLeague//",league_acro,"_",season,"_Run",n_run,"_CMP",".RData"))
  load(file = paste0("Data//MH_Results//Pois_FullLeague//",league_acro,"_",season,"_Run",n_run,"_Pois",".RData"))
  
  t = MH_P$iterations
  home_p = MH_P$home_post[(t/5+1):t]
  home_sas = MH_SAS$home_post[(t/5+1):t]
  home_cmp = MH_CMP$home_post[(t/5+1):t]
  
  p_quants = quantile(home_p, probs = c(0.05, 0.5, 0.95))
  sas_quants = quantile(home_sas, probs = c(0.05, 0.5, 0.95))
  cmp_quants = quantile(home_cmp, probs = c(0.05, 0.5, 0.95))
  
  plot_df = rbind(plot_df,c(p_quants, season, 'Pois'))
  plot_df = rbind(plot_df,c(sas_quants, season, 'CMP-SAS'))
  plot_df = rbind(plot_df,c(cmp_quants, season, 'CMP-Full'))
  
  # cp_toplot = CP_summ[c('home', 'nu_home', 'nu_away'), c('50%', '2.5%', '97.5%')]
  # cp_toplot = cbind(cp_toplot, season)
  # cp_toplot = cbind(cp_toplot, c("CMP-GD", 'CMP_nu_home', 'CMP_nu_away'))
  # plot_df = rbind(plot_df, cp_toplot[,c(4,1,5,2,3)])
  # 
  # p_toplot = c(P_summ[1, c('50%', '2.5%', '97.5%')], season = season, series = 'P')
  # plot_df = rbind(plot_df, p_toplot[c(4,1,5,2,3)])
  # 
  # cpd_toplot = c(CPd_summ[1, c('50%', '2.5%', '97.5%')], season = season, series = 'CMP-ID')
  # plot_df = rbind(plot_df, cpd_toplot[c(4,1,5,2,3)])
}
names(plot_df) = c('lower', 'value', 'upper', 'season', 'series')
plot_df = plot_df[order(plot_df$series),]
row.names(plot_df) <- NULL

plot_df = transform(plot_df, value = as.numeric(value), lower = as.numeric(lower), upper = as.numeric(upper))

# plot_df = plot_df[plot_df$series %in% c('Pois', 'CMP'),]

p = ggplot(plot_df, aes(x = season, y = value, color = series, group = series)) +
  geom_line(linewidth = 1) +                # Line plot
  geom_point(size = 2, alpha = 0.7) +  # Points with size and transparency
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = series),
              alpha = 0.2, linetype = "blank") + # Ribbon for intervals
  theme_minimal() +                    # Minimal theme
  labs(#title = "Estimated parameters (mode) over the seasons with 95% Credible Intervals",
    x = "Season",
    y = "Home Coefficient",
    color = "Series", 
    fill = "Series") +              # Customizing legend title
  theme(legend.position = "bottom")  +
  geom_line(y = 1, linewidth = 1, col = "blue", linetype = 'dashed') +
  geom_line(y = 0, linewidth = 1, col = "blue", linetype = 'dashed')
```

###### Dispersion

```{r}
library(ggplot2)
library(patchwork)
source('utils.R')
library(dplyr)
library(tidyr)
library(forcats)
```

```{r}
league_acro = 'PL'
season = '2324'
n_run = 1
load(file = paste0("Data//MH_Results//CMP_FullLeague//",league_acro,"_",season,"_Run",n_run,"_CMP_P5.RData"))
MH_object = MH_CMP
t = MH_object$iterations
N = length(MH_object$team_names)

nus = MH_object$nu_post[(t/5):t, ]

quantiles <- apply(nus, 2, function(x) quantile(x, probs = c(0.25, 0.75), na.rm = TRUE))

classify <- function(low, high) {
  if (low > 1) {
    return("Underdispersed")
  } else if (high < 1) {
    return("Overdispersed")
  } else {
    return("Equidispersed")
  }
}

group_colors <- c("Underdispersed" = "#D55E00", "Overdispersed" = "#009E73", "Equidispersed" = "gray")

groups <- mapply(classify, quantiles[1, ], quantiles[2, ])
names(groups) <-  MH_object$team_names

df_long <- as.data.frame(nus)
colnames(df_long) <- MH_object$team_names

df_long <- df_long %>%
  pivot_longer(cols = everything(), names_to = "team", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(group = groups[team])

df_long <- df_long %>%
  mutate(group = factor(group, levels = c("Underdispersed", "Equidispersed", "Overdispersed")))

# library(dplyr)

summary_df <- df_long %>%
  group_by(team, group) %>%
  summarise(
    ymin = quantile(log(value), 0.10),
    lower = quantile(log(value), 0.25),
    middle = quantile(log(value), 0.5),
    upper = quantile(log(value), 0.75),
    ymax = quantile(log(value), 0.90),
    .groups = "drop"
  )

# summary_df <- summary_df %>%
#   mutate(team = fct_reorder(team, middle))   # reorder by median
summary_means <- df_long %>%
  group_by(team) %>%
  summarise(mean_val = median(log(value), na.rm = TRUE))

summary_df <- summary_df %>%
  left_join(summary_means, by = "team") %>%
  mutate(team = fct_reorder(team, mean_val))

p1 = ggplot(summary_df, aes(x = team, fill = group)) +
  geom_boxplot(aes(
    ymin = ymin,
    lower = lower,
    middle = middle,
    upper = upper,
    ymax = ymax
  ), stat = "identity", color = "black") +
  scale_fill_manual(values = group_colors) +
  theme_minimal() +
  theme(
    # axis.text.y = element_text(size = 12, family = "sans"),  # team labels
    # axis.text.x = element_text(size = 12, family = "sans"),  # log(v) scale
    # legend.margin = margin(t = -15, b = 0, unit = "pt")
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  ) +
  labs(
    title = '(a)',
    y = expression(logP(nu[i] ~ "|" ~  data)), x = "", fill = "") +
  coord_flip(ylim = c(-2, 0.8)) +
  geom_hline(yintercept = 0, color = "blue", linetype = "dashed", linewidth = 0.7)

# combined_full = p3 + p1 + p2#plot_layout(ncol = 2, widths = c(2, 1.9))

league_acro = 'PL'
season = '2324'
n_run = 1
# load(file = paste0("Data//MH_Results//SAS_FullLeague//",league_acro,"_",season,"_Run",n_run,"_SAS.RData"))
load(file = paste0("Data//MH_Results//SAS_FullLeague//",league_acro,"_",season,"_Prior5_SAS.RData"))
MH_object = MH_SAS
t = MH_object$iterations
N = length(MH_object$team_names)

nus = MH_object$nu_post[(t/5):t, ]
# nus[nus == 1] = NA

quantiles <- apply(nus, 2, function(x) quantile(x, probs = c(0.25, 0.75), na.rm = TRUE))
# colnames(quantiles) = MH_object$team_names

classify <- function(low, high) {
  if (high > 1) {
    return("Underdispersed")
  } else if (low < 1) {
    return("Overdispersed")
  } else {
    return("Equidispersed")
  }
}

group_colors <- c("Underdispersed" = "#D55E00", "Overdispersed" = "#009E73", "Equidispersed" = "gray")

groups <- mapply(classify, quantiles[1, ], quantiles[2, ])
names(groups) <-  MH_object$team_names

df_long <- as.data.frame(nus)
colnames(df_long) <- MH_object$team_names

df_long <- df_long %>%
  pivot_longer(cols = everything(), names_to = "team", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(group = groups[team])

df_long <- df_long %>%
  mutate(group = factor(group, levels = c("Underdispersed", "Equidispersed", "Overdispersed")))

# library(dplyr)

summary_df2 <- df_long %>%
  group_by(team, group) %>%
  summarise(
    ymin = quantile(log(value), 0.10),
    lower = quantile(log(value), 0.25),
    middle = quantile(log(value), 0.5),
    upper = quantile(log(value), 0.75),
    ymax = quantile(log(value), 0.90),
    .groups = "drop"
  )

summary_df2 <- summary_df2 %>%
  mutate(team = factor(team, levels = levels(summary_df$team))) %>%
  arrange(team)

p2 = ggplot(summary_df2, aes(x = team, fill = group)) +
  geom_boxplot(aes(
    ymin = ymin,
    lower = lower,
    middle = middle,
    upper = upper,
    ymax = ymax
  ), stat = "identity", color = "black") +
  scale_fill_manual(values = group_colors) +
  theme_minimal() +
  theme(
    # axis.text.y = element_text(size = 12, family = "sans"),  # team labels
    # axis.text.x = element_text(size = 12, family = "sans"),  # log(v) scale
    # legend.margin = margin(t = -15, b = 0, unit = "pt")
    axis.text.y = element_blank(),
    plot.title = element_text(hjust = 0.5),
    legend.position = "none"
  ) +
  labs(
    title = "(b)",
    y = expression(logP(nu[i] ~ "|" ~  data)), x = "", fill = "") +
  coord_flip(ylim = c(-2, 0.8)) +
  geom_hline(yintercept = 0, color = "blue", linetype = "dashed", linewidth = 0.7)


p_z1 = retrieve_nu_sas_summ(MH_SAS$Z_post, MH_SAS$nu_post, (t/5)+1, t)[1,]
names(p_z1) = MH_SAS$team_names

team_levels <- levels(summary_df$team)
p_z1 <- p_z1[team_levels]

df_z1 <- data.frame(
  team = factor(names(p_z1), levels = levels(summary_df$team)),
  p_z1 = as.numeric(p_z1)
)

p3 = ggplot(df_z1, aes(x = p_z1, y = team)) +
  geom_col(fill = "skyblue3", width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", p_z1)),
            hjust = -0.1, size = 3.3) +
  scale_x_continuous(limits = c(0, 1)) +
  labs(
    title = "(c)",
    x = expression(P(Z[i] == 1 ~ "|" ~ data)),
    y = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),   # removes team names
    axis.ticks.y = element_blank(),  # removes ticks
    panel.grid.minor = element_blank(),
    axis.title.x = element_text(margin = margin(b = 2)),
    plot.title = element_text(hjust = 0.5)
  )

combined_plot <- p1 + p2 + p3
# combined_plot <- p1 + p2 + plot_layout(ncol = 2, widths = c(2, 1.9))
```

####### Over Seasons

```{r}
# Required packages
library(tidyverse)
library(ggplot2)
# library(ggpattern)
league_acro = 'SA'
# league_acro = 'SA'
# season = '2425'
team_to_plot = 'Juventus'

seasons_strvec = generate_season_string(2015, 2025)

nus_bigmatrix = matrix(nrow = 40000, ncol = 10)
colnames(nus_bigmatrix) = seasons_strvec

n_run = 1
for (season in seasons_strvec){
  load(file = paste0("Data//MH_Results//SAS_FullLeague//",league_acro,"_",season,"_Run",n_run,"_SAS_CC20",".RData"))
  
  MH_object = MH_SAS
  t = MH_object$iterations
  N = length(MH_object$team_names)
  
  # pz1 = retrieve_nu_sas_summ(MH_SAS$Z_post, MH_SAS$nu_post, 10001, 50000)[1,]
  # filter_idx = which(pz1 > threshold)
  #     team_filter = MH_SAS$team_names[filter_idx]
  #     print(team_filter)
  #     print(pz1[filter_idx])

  if (team_to_plot %in% MH_object$team_names){
      team_idx = which(MH_object$team_names == team_to_plot)

  nus_bigmatrix[,season] = MH_object$nu_post[(t/5+1):t, team_idx]
  }
  else{
    next
  }
}

# nus_bigmatrix[nus_bigmatrix == 1] = NA

# library(tidyverse)

# Assume param_matrix is your matrix (rows = samples, cols = seasons)
# Take the log
param_log <- log(nus_bigmatrix)

# Convert to long format
df_long <- as.data.frame(param_log) %>%
  mutate(id = row_number()) %>%
  pivot_longer(-id, names_to = "season", values_to = "value") %>%
  filter(!is.na(value))

# Compute IQR and classify each season
summary_df <- df_long %>%
  group_by(season) %>%
  summarise(
    ymin   = quantile(value, 0.10, na.rm = TRUE),
    lower  = quantile(value, 0.25, na.rm = TRUE),
    middle = quantile(value, 0.50, na.rm = TRUE),
    upper  = quantile(value, 0.75, na.rm = TRUE),
    ymax   = quantile(value, 0.90, na.rm = TRUE)
  ) %>%
  mutate(
    group = case_when(
      upper > 0 ~ "Underdispersed",
      lower < 0 ~ "Overdispersed",
      TRUE ~ "Equidispersed"
    )
    #,
    # color = case_when(
    #   group == "Overdispersed" ~ "#009E73",      # green
    #    group == "Equidispersed" ~ "gray",
    #   group == "Underdispersed" ~ "#D55E00")
  )

group_colors <- c("Underdispersed" = "#D55E00", "Overdispersed" = "#009E73", "Equidispersed" = "gray")

summary_df <- summary_df %>%
  mutate(group = factor(group, levels = c("Underdispersed", "Equidispersed", "Overdispersed")))
# Plot using stat = "identity"
p = ggplot(summary_df, aes(x = season)) +
  geom_boxplot(
    aes(
      ymin = ymin, lower = lower, middle = middle,
      upper = upper, ymax = ymax, fill = group
    ),
    stat = "identity",
    color = "black",
    width = 0.6
  ) +
  scale_fill_manual(
    values = group_colors,
    name = ""
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "blue") +
  theme_minimal() +
  labs(
    x = "Season",
    y =  expression(logP(nu[JUV] * "|" * data)),
    title = ""
  ) +
  theme(
    axis.text.x = element_text(angle = 50, vjust = 0.5, hjust = 0.5, size = 14),
    plot.title = element_text(face = "bold"),
    legend.margin = margin(t = -15, b = 0, unit = "pt") 
    # legend.position = c(0.90, 0.3)  # adjust as needed
  )
```

####### Over Games in a single season

```{r}
start_year = 2015
end_year = 2025
seasons_strvec = generate_season_string(start_year, end_year)

n_seeds = 1

league_acros = c('PL', 'SA', 'LL', 'LC', 'BL')

leagues = c('Premier', 'SerieA', 'Liga', 'Ligue', 'Bundes')

L = 1
league = leagues[L]
league_acro = league_acros[L]
print(league)
print(league_acro)

get_summary <- function(mat, team_name) {
  summary_stats <- apply(mat, 2, function(x) {
    c(
      median = median(x, na.rm = TRUE),
      lower = quantile(x, 0.25, na.rm = TRUE),
      upper = quantile(x, 0.75, na.rm = TRUE)
    )
  })

  summary_df <- as.data.frame(t(summary_stats))
  colnames(summary_df) <- c("median", "lower", "upper")  # explicitly rename columns

  summary_df$time <- game_n
  summary_df$team <- team_name

  return(summary_df)
}

season = '2425'
nus_df = NULL
games_seq = 10*seq(19,37, by=1)
    for (n_games in games_seq){
      load(file = paste0("Data//MH_Results//SAS_MidLeague//",league_acro,"_",season,"_n",n_games,"_SAS",".RData"))
      MH_object = MH_SAS
      t = MH_object$iterations
      team_names = MH_object$team_names
      nus = MH_object$nu_post[(t/2):t,]
      colnames(nus) = team_names
      teams_selected = c('Man United', 'Liverpool')
      nus = nus[,teams_selected]
      # nus = MH_object$nu_post[(t/2):t, which(team_names %in% c('Atalanta', 'Milan'))]
      nus[nus == 1] = NA
      nus = log(nus)
      
      nus_df = rbind(nus_df, 
                     nus %>%
                          apply(2, function(x){
                            c(
                              median = median(x, na.rm = TRUE),
                              lower = quantile(x, 0.10, na.rm = TRUE),
                              upper = quantile(x, 0.90, na.rm = TRUE)
                            )
                          }) %>%
                          t() %>%
                          as.data.frame() %>%
                          mutate(game_n = n_games, team = teams_selected)
                    )
    }

colnames(nus_df) = c('median', 'lower', 'upper', 'game_n', 'team')

team_colors <- c("Man United" = "#00539F", "Liverpool" = "#DA291C")  # example
desired_order <- c("Liverpool", "Man United")  # or any order you prefer
p = ggplot(nus_df, aes(x = game_n, y = median, group = team, color = team, fill = team)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, color = NA) +
  geom_line(size = 1) +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'blue') +
  theme_minimal() +
  labs(
    x = "No. of games",
    y = "log(v)",
    color = "Team",  # legend title for lines
    fill = "Team"    # legend title for ribbons
  ) +
  # scale_x_discrete(labels = games_seq, drop = FALSE) +
  # theme(
  #   axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
  #   axis.ticks.x = element_line(color = "black"),
  #   axis.line.x = element_line(color = "black"),
  #   panel.grid.minor.x = element_blank(),
  #   panel.grid.major.x = element_blank()
  # ) +
  scale_color_manual(
    values = team_colors,  # define this named vector below
    breaks = desired_order  # control legend order
  ) +
  scale_fill_manual(
    values = team_colors,
    breaks = desired_order
  )
p
```

###### Posterior Predictive Plot

```{r}
library(reshape2)
library(patchwork)
library(dplyr)
library(tidyr)

rotate_ccw <- function(mat) {
  t(apply(mat, 2, rev))
}

plot_heatmap <- function(mat, title = "", x = NULL, y = NULL) {
  df <- reshape2::melt((mat))
  colnames(df) <- c("Away", "Home", "Odd")  # match orientation

  df$Home <- df$Home - 1
  df$Away <- df$Away - 1

  p <- ggplot(df, aes(x = Home, y = Away, fill = Odd)) +
    geom_tile(color = "white") +
    geom_text(aes(label = round(Odd, 2)), size = 4, color = "black") +
    scale_fill_gradient(low = "gray95", high = "red4", guide = 'none') +
    scale_x_continuous(breaks = 0:7, name = "Home Goals") +
    scale_y_continuous(breaks = 0:7, name = "Away Goals") +
    coord_fixed(ratio = 0.7) +  # <- Make cells wider
    labs(title = title) +
    theme_minimal(base_size = 12)

  # Add border for selected cell if x and y are provided
  if (!is.null(x) && !is.null(y)) {
    p <- p + geom_tile(
      data = subset(df, Home == x & Away == y),
      aes(x = Home, y = Away),
      fill = NA, color = "black", size = 1.2
    )
  }

  return(p)
}

```

```{r}
league_acro = 'PL'
season = '2324'
df_hist = read_historics(season, league_acro)
game_id = 254

row_hist = df_hist[game_id, ]
ht = row_hist[['HomeTeam']]
at = row_hist[['AwayTeam']]
hg = row_hist[['FTHG']]
ag = row_hist[['FTAG']]

load(file = paste0("Data//Predictions//",league_acro,season,"_SAS_oos.RData"))
sas_preds = predictions

cmp_mat = sas_preds$list_preds[[game_id]]

cmp_heatplot = plot_heatmap(cmp_mat, "CMP-SAS Model", x = as.numeric(hg), y = as.numeric(ag))


load(file = paste0("Data//Predictions//",league_acro,season,"_Pois_oos.RData"))
pois_preds = predictions

pois_mat = pois_preds$list_preds[[game_id]]
pois_heatplot = plot_heatmap(pois_mat, "Poisson Model",x = as.numeric(hg), y = as.numeric(ag))

combined_plot = (pois_heatplot | plot_spacer() |cmp_heatplot) +
  plot_layout(widths = c(1, 0, 1)) +
  plot_annotation(title = paste0("(Home) ",ht,"-",at," (Away), Result: ",hg,"-",ag, ", Premier League ", season, ", Out of sample prediction"))

# combined_plot = (pois_heatplot + cmp_heatplot) +
#   plot_layout(widths = c(1, 1)) +
#   plot_annotation(title = "(Home)Man City-Liverpool(Away), Result: 0-2, Premier League 2024/25")

cmp_prob = c(sum(cmp_mat[upper.tri(cmp_mat)]),
             sum(diag(cmp_mat)),
             sum(cmp_mat[lower.tri(cmp_mat)]))
pois_prob = c(sum(pois_mat[upper.tri(pois_mat)]),
              sum(diag(pois_mat)),
              sum(pois_mat[lower.tri(pois_mat)]))

print(cmp_prob)
print(pois_prob)
ftr_map <- c(H = 1, D = 2, A = 3)
outcome = ftr_map[[ df_hist[game_id,'FTR'] ]]
outcome_rps(cmp_prob/100, outcome)
outcome_rps(pois_prob/100, outcome)
```

```{r}
cmp_prob = c(sum(cmp_mat[upper.tri(cmp_mat)]),
             sum(diag(cmp_mat)),
             sum(cmp_mat[lower.tri(cmp_mat)]))
pois_prob = c(sum(pois_mat[upper.tri(pois_mat)]),
              sum(diag(pois_mat)),
              sum(pois_mat[lower.tri(pois_mat)]))
exact_out = c(0,0,1)

x_vals = 4:-4
gd_outcomes = 4:-4
hg = min(as.numeric(df_hist[game_id, 'FTHG']), 7)
      ag = min(as.numeric(df_hist[game_id, 'FTAG']), 7)
      goal_diff_obs = hg - ag
      
exact_obs = as.integer(gd_outcomes == goal_diff_obs)
names(exact_obs) = x_vals
cum_gd_obs = as.integer(gd_outcomes <= goal_diff_obs)

pois_goaldiff = (compute_goaldiff_probs(pois_mat)/100)[4:12]
cmp_goaldiff = (compute_goaldiff_probs(cmp_mat)/100)[4:12]
      
cum_pois = cumsum(pois_goaldiff)
cum_sas = cumsum(cmp_goaldiff)

categories <- factor(c("Home Win", "Draw", "Away Win"),
                     levels = c("Home Win", "Draw", "Away Win"))

# --------------------- Shared color scale ------------------------
my_colors <- c("pois" = "#1f77b4", "cmp" = "#ff7f0e", "Realized outcome" = "blue4")
my_labels <- c("pois" = "Pois", "cmp" = "CMP-SAS", "Realized outcome" = "Realized outcome")
my_scale <- scale_color_manual(values = my_colors, labels = my_labels, name = NULL)

# --------------------- p1: predictive densities (match outcome) --------------
df1 <- data.frame(
  category = categories,
  pois = pois_prob/100,
  cmp  = cmp_prob/100,
  observed = exact_out
) %>%
  pivot_longer(cols = c(pois, cmp), names_to = "model", values_to = "prob")

obs_category <- "Away Win"
df1_obs <- data.frame(category = obs_category, model = "Realized outcome")

p1 <- ggplot(df1, aes(x = category, y = prob, group = model)) +
  geom_line(aes(color = model), size = 1) +
  geom_point(aes(color = model), size = 2) +
  geom_vline(data = df1_obs,
             aes(xintercept = category, color = model),
             linetype = "dashed", linewidth = 1, alpha = 0.4) +
  my_scale +
  labs(y = "") +
  theme_minimal() +
  theme(legend.position = "bottom",
        axis.title.x = element_blank(),
        axis.title.y = element_text())

# --------------------- p2: predictive densities (goal difference) --------------
goal_diff_vals <- as.numeric(names(pois_goaldiff))
df2 <- data.frame(
  goal_diff = goal_diff_vals,
  pois = as.numeric(pois_goaldiff),
  cmp  = as.numeric(cmp_goaldiff),
  observed = as.numeric(exact_obs)
) %>%
  pivot_longer(cols = c(pois, cmp), names_to = "model", values_to = "prob") %>%
  filter(prob > 0)

obs_goaldiff <- -1
df2_obs <- data.frame(x = obs_goaldiff, model = "Realized outcome")

p2 <- ggplot(df2, aes(x = goal_diff, y = prob, group = model)) +
  geom_line(aes(color = model), size = 1) +
  geom_point(aes(color = model), size = 2) +
  geom_vline(data = df2_obs,
             aes(xintercept = x, color = model),
             linetype = "dashed", linewidth = 1, alpha = 0.4) +
  my_scale +
  scale_x_reverse(breaks = sort(unique(df2$goal_diff), decreasing = TRUE)) +
  labs(y = "", x = "Goal difference") +
  theme_minimal() +
  theme(legend.position = "bottom")

combined_plot_1 = p1 + p2 + plot_layout(guides = "collect") +
  theme(legend.position = "bottom")

# --------------------- p3: cumulative densities (match outcome) --------------
df1_cum <- df1 %>%
  group_by(model) %>%
  arrange(category) %>%
  mutate(cum_prob = cumsum(prob) / sum(prob))

df1_obs_cum <- data.frame(
  category = levels(df1$category),
  cum_prob = as.numeric(seq_along(levels(df1$category)) >= which(levels(df1$category) == obs_category)),
  model = "Realized outcome"
)

p3 <- ggplot() +
  geom_line(data = df1_cum, aes(x = category, y = cum_prob, color = model, group = model), size = 1) +
  geom_point(data = df1_cum, aes(x = category, y = cum_prob, color = model, group = model), size = 2) +
  geom_step(data = df1_obs_cum, aes(x = category, y = cum_prob, color = model, group = model),
            linetype = "dashed", size = 1, alpha = 0.5) +
  my_scale +
  labs(y = "Cumulative probability") +
  theme_minimal() +
  theme(legend.position = "bottom",
        axis.title.x = element_blank(),
        axis.title.y = element_text())

# --------------------- p4: cumulative densities (goal difference) --------------
df2_cum <- df2 %>%
  group_by(model) %>%
  arrange(desc(goal_diff)) %>%  # reverse order for accumulation
  mutate(cum_prob = cumsum(prob)) %>%
  arrange(goal_diff)

df2_obs_cum <- data.frame(
  goal_diff = goal_diff_vals,
  cum_prob = as.numeric(goal_diff_vals <= obs_goaldiff),
  model = "Realized outcome"
) %>%
  arrange(desc(goal_diff))

p4 <- ggplot() +
  geom_line(data = df2_cum, aes(x = goal_diff, y = cum_prob, color = model, group = model), size = 1) +
  geom_point(data = df2_cum, aes(x = goal_diff, y = cum_prob, color = model, group = model), size = 2) +
  geom_step(data = df2_obs_cum, aes(x = goal_diff, y = cum_prob, color = model, group = model),
            linetype = "dashed", size = 1, alpha = 0.5) +
  my_scale +
  scale_x_reverse()+
  labs(y = "", x = "Goal difference") +
  theme_minimal() +
  theme(legend.position = "bottom")

# --------------------- Combine 4 plots with single legend --------------------
combined_plot <- (p1 + p2) / (p3 + p4) +
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

```

```{r}
library(patchwork)
n_samples = 5000
# df_odds = read.csv(file = paste0("Data/Historics/",league_acro,"_Odds_",season,".csv"))
set.seed(1)
# n_games = 190
ids = get_sample_ids(n_samples, t, t/5)
# data.frame(home_goals,away_goals)
df_hist = read_historics('2324', 'PL')

load(file = paste0("Data//MH_Results//SAS_FullLeague//",league_acro,"_",season,"_Run",n_run,"_SAS_CC20",".RData"))
# MH_object = MH_SAS
MH_object = thin_mcmc(MH_SAS, 10)
att_samples = MH_object$att_post[ids,]
def_samples = MH_object$def_post[ids,]
nu_samples = MH_object$nu_post[ids,]
home_samples = MH_object$home_post[ids]
m = 100
team_names = MH_object$team_names
team_codes = setNames(1:N, team_names)
for (game_id in 378){
      i = team_codes[df_hist[game_id, 'HomeTeam']]
      j = team_codes[df_hist[game_id, 'AwayTeam']]
      
      mu_home = exp(att_samples[,i] + def_samples[,j] + home_samples)
      nu_home = nu_samples[,i]
      mu_away = exp(att_samples[,j] + def_samples[,i])
      nu_away = nu_samples[,j]
      
      home_goals = matrix(NA_real_, nrow = n_samples, ncol = m)
      away_goals = matrix(NA_real_, nrow = n_samples, ncol = m)
      for (s in 1:n_samples){
        home_goals[s,] = rejection_sampler_draws(m, mu_home[s], nu_home[s])
        away_goals[s,] = rejection_sampler_draws(m, mu_away[s], nu_away[s])
      }
}
cmp_mat = create_heatmatrix(c(home_goals), c(away_goals))[[2]]/(n_samples*m/100)
cmp_heatplot = plot_heatmap(cmp_mat, "CMP-SAS Model", x = 0, y = 2)
cmp_heatplot

load(file = paste0("Data//MH_Results//Pois_FullLeague//",league_acro,"_",season,"_Run",n_run,"_Pois_CC20",".RData"))
MH_object = MH_P
att_samples = MH_object$att_post[ids,]
def_samples = MH_object$def_post[ids,]
nu_samples = MH_object$nu_post[ids,]
home_samples = MH_object$home_post[ids]

for (game_id in 378){
  # game_id = 189
  i = team_codes[df_hist[game_id, 'HomeTeam']]
  j = team_codes[df_hist[game_id, 'AwayTeam']]
  
  mu_home = exp(att_samples[,i] + def_samples[,j] + home_samples)
  nu_home = nu_samples[,i]
  mu_away = exp(att_samples[,j] + def_samples[,i])
  nu_away = nu_samples[,j]
  
  home_goals = matrix(NA_real_, nrow = n_samples, ncol = m)
  away_goals = matrix(NA_real_, nrow = n_samples, ncol = m)
  for (s in 1:n_samples){
    home_goals[s,] = rejection_sampler_draws(m, mu_home[s], 1)
    away_goals[s,] = rejection_sampler_draws(m, mu_away[s], 1)
    # home_goals[s,] = rpois(m, mu_home[s])
    # away_goals[s,] = rpois(m, mu_away[s])
  }
  
}

pois_mat = create_heatmatrix(c(home_goals), c(away_goals))[[2]]/(n_samples*m/100)
pois_heatplot = plot_heatmap(pois_mat, "Poisson Model", x = 0, y = 2)

combined_plot = (pois_heatplot | plot_spacer() |cmp_heatplot) +
  plot_layout(widths = c(1, 0, 1)) +
  plot_annotation(title = "(Home)Man City-Liverpool(Away), Result: 0-2, Premier League 2024/25")

combined_plot = (pois_heatplot + cmp_heatplot) +
  plot_layout(widths = c(1, 1)) +
  plot_annotation(title = "(Home)Man City-Liverpool(Away), Result: 0-2, Premier League 2024/25")
sum(diag(pois_mat))
sum(diag(cmp_mat))
sum(pois_mat[upper.tri(pois_mat)])
sum(cmp_mat[upper.tri(cmp_mat)])

sum(pois_mat[lower.tri(pois_mat)])
sum(cmp_mat[lower.tri(cmp_mat)])

create_heatmatrix_fast <- function(home, away) {
  nmax <- 8
  
  # Clip values to max 7 (as 0-based index), then add 1 for 1-based R indexing
  i <- pmin(home, nmax - 1) + 1
  j <- pmin(away, nmax - 1) + 1
  
  # Use table to count occurrences efficiently
  heatmatrix <- matrix(0, nrow = nmax, ncol = nmax)
  tab <- table(factor(i, levels = 1:nmax), factor(j, levels = 1:nmax))
  heatmatrix[as.matrix(expand.grid(1:nmax, 1:nmax)[, 2:1])] <- as.vector(tab)
  
  # Melt with row/col labels starting from 0
  m.hmat <- reshape2::melt(heatmatrix)
  m.hmat[, 1:2] <- m.hmat[, 1:2] - 1
  
  return(list(m.hmat, heatmatrix))
}
fast_mat= create_heatmatrix_fast(c(home_goals), c(away_goals))[[2]]/5000
```

```{r}
sim_halfleague_results_pois = function(season, league_acro, m = 10, n_samples = 5000){
  df_hist = read_historics(season, league_acro)
  team_names = sort(unique(df_hist[,"HomeTeam"]))
  N = length(team_names)
  team_codes = setNames(1:N, team_names)
  tot_matches = nrow(df_hist)
  
  list_preds = vector("list", length = tot_matches/2)
  games_seq = 10*seq(19,37, by=1)
  for (n_games in games_seq){
    load(file = paste0("Data//MH_Results//Pois_MidLeague//",league_acro,"_",season,"_n",n_games,"_Pois",".RData"))
    
    MH_object = MH_SAS
    
    t = MH_object$iterations
    
    ids = get_sample_ids(n_samples, t, t/2)
    
    att_samples = MH_object$att_post[ids,]
    def_samples = MH_object$def_post[ids,]
    # nu_samples = MH_object$nu_post[ids,]
    home_samples = MH_object$home_post[ids]
    
    for (game_id in (n_games+1):(n_games+10)){
      # game_id = 189
      i = team_codes[df_hist[game_id, 'HomeTeam']]
      j = team_codes[df_hist[game_id, 'AwayTeam']]
      
      print(paste0(i, ' vs ', j, '; game-id: ', game_id))
      
      mu_home = exp(att_samples[,i] + def_samples[,j] + home_samples)
      # nu_home = nu_samples[,i]
      mu_away = exp(att_samples[,j] + def_samples[,i])
      # nu_away = nu_samples[,j]
      
      home_goals = matrix(NA_real_, nrow = n_samples, ncol = m)
      away_goals = matrix(NA_real_, nrow = n_samples, ncol = m)
      for (s in 1:n_samples){
        # home_goals[s,] = rejection_sampler_draws(m, mu_home[s], nu_home[s])
        # away_goals[s,] = rejection_sampler_draws(m, mu_away[s], nu_away[s])
        home_goals[s,] = rpois(m, mu_home[s])
        away_goals[s,] = rpois(m, mu_away[s])
      }
      
      list_preds[[game_id - 190]] = create_heatmatrix(c(home_goals), c(away_goals))[[2]]/5000  
    }
  }
  return(list_preds)
}
```

```{r}
pois_preds = sim_halfleague_results_pois(season, league_acro, m = 1)
```

#### Simulation SAS Plots

```{r}
nus_overdispersed = seq(0.2, 0.90, by = 0.1)
n_seeds = 5
threshold = 0.5

big_Z_nd = matrix(NA, nrow = 50, ncol = length(nus_overdispersed))
big_Z_ed = matrix(NA, nrow = 50, ncol = length(nus_overdispersed))

significant_nu_nd = matrix(NA, nrow = 50, ncol = length(nus_overdispersed))
significant_nu_ed = matrix(NA, nrow = 50, ncol = length(nus_overdispersed))
  
colnames(big_Z_nd) = nus_overdispersed
colnames(big_Z_ed) = nus_overdispersed

colnames(significant_nu_nd) = nus_overdispersed
colnames(significant_nu_ed) = nus_overdispersed

for (nu_scalar in nus_overdispersed){
  load(file = paste0("Data//Simulations//SAS_sim_nubase_",gsub("\\.", "_", as.character(nu_scalar)),".RData"))
  vec_Z_nd = NULL
  vec_Z_ed = NULL
  
  vec_sign_nu_nd = NULL
  vec_sign_nu_ed = NULL
  for (run in 1:n_seeds){
    sim = simulations_list[[run]]
    
    load(file = paste0("Data//MH_Results//Simulations/SAS_SIM_Nu",gsub("\\.", "_", as.character(nu_scalar)),"_Run",run,".RData"))
    t = MH_SAS$iterations
    
    Z_nd = MH_SAS$Z_post[(t/5):t, sim$nu_true != 1 ]
    Z_ed = MH_SAS$Z_post[(t/5):t, sim$nu_true == 1]
    vec_Z_nd = c(vec_Z_nd, colMeans(Z_nd))
    vec_Z_ed = c(vec_Z_ed, colMeans(Z_ed))
    
    {
      nu_summ = retrieve_nu_sas_summ(MH_SAS$Z_post, MH_SAS$nu_post, t/5, t, sim$nu_true)
      # included <- as.integer(!((nu_summ[3, ] <= 1) & (1 <= nu_summ[5, ])))
      included = as.integer(nu_summ[1,] >= threshold)
      
      vec_sign_nu_nd = c(vec_sign_nu_nd, included[sim$nu_true != 1])
      vec_sign_nu_ed = c(vec_sign_nu_ed, included[sim$nu_true == 1])
    }
  }  
  
  big_Z_nd[,as.character(nu_scalar)] = vec_Z_nd
  big_Z_ed[,as.character(nu_scalar)] = vec_Z_ed
  
  significant_nu_nd[,as.character(nu_scalar)] = vec_sign_nu_nd
  significant_nu_ed[,as.character(nu_scalar)] = vec_sign_nu_ed
}
```

```{r}
df1 <- as.data.frame(big_Z_nd)
df2 <- as.data.frame(big_Z_ed)

df1$ID <- 1:nrow(df1)
df2$ID <- 1:nrow(df2)
df1$Group <- "Overdispersed"
df2$Group <- "Equidispersed"
df_long <- bind_rows(df1, df2)

df_long <- pivot_longer(
  df_long,
  cols = all_of(as.character(seq(0.2, 0.9, by = 0.1))),
  names_to = "Variable",
  values_to = "Value"
)

p = ggplot(df_long, aes(x = Group, y = Value, fill = Group)) +
  geom_boxplot(outlier.size = 0.8, width = 0.6) +
  geom_hline(yintercept = 0.5, linetype = 'dashed', linewidth = 0.5, color = 'red') +
  facet_wrap(~ Variable, nrow = 2) +  # Adjust nrow/ncol as needed
  coord_cartesian(ylim = c(0, 1)) + 
  theme_minimal() +
  scale_fill_manual(
    name = "",  # Legend title
    values = c("Overdispersed" = "#1b9e77", "Equidispersed" = "#1f77b4")  # Custom colors
  ) +
  theme(
    axis.text.x = element_blank(),
    legend.position = "bottom",
    legend.box.margin = margin(t = -10),
    legend.margin = margin(t = -5),
    strip.text = element_text(face = "bold")
  ) +
  labs(title = "", y = expression(E * "(" * P(Z[i] == 1 ~ "|" ~ bold(Y)) * ")"), x = "")

```

```{r}
df1 <- as.data.frame(big_Z_nd)
df2 <- as.data.frame(big_Z_ed)
df1$ID <- 1:nrow(df1)
df2$ID <- 1:nrow(df2)
df1$Group <- "Overdispersed"
df2$Group <- "Equidispersed"
df_long <- bind_rows(df1, df2)

df_long <- pivot_longer(
  df_long,
  cols = all_of(as.character(seq(0.1, 0.90, by = 0.1))),
  names_to = "Variable",
  values_to = "Value"
)

significancy_vector = c(as.vector(t(significant_nu_nd)), as.vector(t(significant_nu_ed)))
df_long = cbind(df_long, Significant = significancy_vector)

df_long <- df_long %>%
 mutate(Significant = factor(Significant, levels = c(0, 1), labels = c("Not Significant", "Significant"))) 

p = ggplot(df_long, aes(x = Group, y = Value, color = Significant)) +
  geom_jitter(
    width = 0.2,
    height = 0,
    size = 1.5,
    alpha = 0.7
  ) +
  facet_wrap(~ Variable, nrow = 3) +
  theme_minimal() +
  theme(
    # axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom"
  ) +
  scale_color_manual(
    values = c("Significant" = "#1b9e77", "Not Significant" = "#d95f02"),
    name = "Value"
  ) +
  labs(
    title = "Jittered Scatterplots by nu_true",
    y = "P(Z=1|data)",
    x = ""
  )

```

```{r}
nus_underdispersed = seq(1.2, 4, by = 0.4)

big_Z_nd = matrix(NA, nrow = 50, ncol = length(nus_underdispersed))
big_Z_ed = matrix(NA, nrow = 50, ncol = length(nus_underdispersed))

significant_nu_nd = matrix(NA, nrow = 50, ncol = length(nus_underdispersed))
significant_nu_ed = matrix(NA, nrow = 50, ncol = length(nus_underdispersed))

colnames(big_Z_nd) = nus_underdispersed
colnames(big_Z_ed) = nus_underdispersed

colnames(significant_nu_nd) = nus_underdispersed
colnames(significant_nu_ed) = nus_underdispersed

n_seeds = 5
for (nu_scalar in nus_underdispersed){
  load(file = paste0("Data//Simulations//SAS_sim_nubase_",gsub("\\.", "_", as.character(nu_scalar)),".RData"))
  vec_Z_nd = NULL
  vec_Z_ed = NULL
  
  vec_sign_nu_nd = NULL
  vec_sign_nu_ed = NULL
  for (run in 1:n_seeds){
    sim = simulations_list[[run]]
    
    load(file = paste0("Data//MH_Results//Simulations/SAS_SIM_Nu",gsub("\\.", "_", as.character(nu_scalar)),"_Run",run,".RData"))
    t = MH_SAS$iterations
    
    Z_nd = MH_SAS$Z_post[(t/5):t, sim$nu_true != 1 ]
    Z_ed = MH_SAS$Z_post[(t/5):t, sim$nu_true == 1]
    vec_Z_nd = c(vec_Z_nd, colMeans(Z_nd))
    vec_Z_ed = c(vec_Z_ed, colMeans(Z_ed))
    
    {
      nu_summ = retrieve_nu_sas_summ(MH_SAS$Z_post, MH_SAS$nu_post, t/5, t, sim$nu_true)
      # included <- as.integer(!((nu_summ[3, ] <= 1) & (1 <= nu_summ[5, ])))
      included = as.integer(nu_summ[1,] >= threshold)
      
      vec_sign_nu_nd = c(vec_sign_nu_nd, included[sim$nu_true != 1])
      vec_sign_nu_ed = c(vec_sign_nu_ed, included[sim$nu_true == 1])
    }
  }  
  
  big_Z_nd[,as.character(nu_scalar)] = vec_Z_nd
  big_Z_ed[,as.character(nu_scalar)] = vec_Z_ed
  
  significant_nu_nd[,as.character(nu_scalar)] = vec_sign_nu_nd
  significant_nu_ed[,as.character(nu_scalar)] = vec_sign_nu_ed
}
```

```{r}
df1 <- as.data.frame(big_Z_nd)
df2 <- as.data.frame(big_Z_ed)

df1$ID <- 1:nrow(df1)
df2$ID <- 1:nrow(df2)
df1$Group <- "Underdispersed"
df2$Group <- "Equidispersed"
df_long <- bind_rows(df1, df2)

df_long <- pivot_longer(
  df_long,
  cols = all_of(as.character(nus_underdispersed)),
  names_to = "Variable",
  values_to = "Value"
)

```

```{r}
p = ggplot(df_long, aes(x = Group, y = Value, fill = Group)) +
  geom_boxplot(outlier.size = 0.8, width = 0.6) +
  geom_hline(yintercept = 0.5, linetype = 'dashed', linewidth = 0.5, color = 'red') +
  facet_wrap(~ Variable, nrow = 2) +  # Adjust nrow/ncol as needed
  coord_cartesian(ylim = c(0, 1)) + 
  theme_minimal() +
  scale_fill_manual(
    name = "",  # Legend title
    values = c("Underdispersed" = "#D55E00", "Equidispersed" = "#1f77b4")  # Custom colors
  ) +
  theme(
    axis.text.x = element_blank(),
    legend.position = "bottom",
    legend.box.margin = margin(t = -10),
    legend.margin = margin(t = -5),
    strip.text = element_text(face = "bold")
  ) +
  labs(title = "", y = expression(E * "(" * P(Z[i] == 1 ~ "|" ~ bold(Y)) * ")"), x = "")
```

```{r}
significancy_vector = c(as.vector(t(significant_nu_nd)), as.vector(t(significant_nu_ed)))
df_long = cbind(df_long, Significant = significancy_vector)

df_summary <- df_long %>%
  group_by(Variable, Group) %>%
  summarise(mean_significance = mean(Significant), .groups = "drop")

df_wide <- df_summary %>%
  pivot_wider(names_from = Variable, values_from = mean_significance)

library(knitr)
library(kableExtra)
# Export as LaTeX table
kable(df_wide, format = "latex", booktabs = TRUE, digits = 3,
      caption = "Mean Significance by Group and Variable") %>%
  kable_styling(latex_options = "hold_position") %>%
  row_spec(0, bold = TRUE)  # 0 means header row

df_long <- df_long %>%
  mutate(Significant = factor(Significant, levels = c(0, 1), labels = c("Not Significant", "Significant")))       



p2 = ggplot(df_long, aes(x = Group, y = Value, color = Significant)) +
  geom_jitter(
    width = 0.2,
    height = 0,
    size = 1.5,
    alpha = 0.7
  ) +
  facet_wrap(~ Variable, nrow = 3) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
   strip.text = element_text(face = "bold")
  ) +
  scale_color_manual(
    values = c("Significant" = "#d73027", "Not Significant" = "#1a9850"),
    name = "Value"
  ) +
  labs(
    title = "Jittered Scatterplots by nu_true",
    y = "P(Z=1|data)",
    x = ""
  )
```

#### Varying N

```{r}
N_vec = c(6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 40, 50, 60)

n_seeds = 3

df_long = data.frame(matrix(ncol = 4, nrow = 0))
colnames(df_long) = c("Group", "Variable", "Value", "Significant")

for (N in N_vec){
  load(file = paste0("Data//Simulations//SAS_sim_N",N,".RData"))
  n_disp = floor(N/3)+1
  n_eq = N-2*n_disp
  for (run in 1:n_seeds){
    load(file = paste0("Data//MH_Results//Simulations//SAS_Simulations/VaryN_4/SAS_SIM_VaryN",N,"_Run",run,".RData"))
    sim = simulations_list[[run]]
    
    idx_ud = sim$nu_true > 1
    idx_eq = sim$nu_true == 1
    idx_od = sim$nu_true < 1
    
    t = MH_SAS$iterations
    
    nu_summ = retrieve_nu_sas_summ(MH_SAS$Z_post, MH_SAS$nu_post, t/2, t, sim$nu_true)
    significant_binary <- as.integer(!((nu_summ[3, ] <= 1) & (1 <= nu_summ[5, ])))
    
    Z_ud = colMeans(MH_SAS$Z_post[(t/2):t, idx_ud])
    df_ud = cbind(Group = rep("Underdispersed", n_disp), 
                  Variable = rep(N, n_disp), 
                  Value = Z_ud, 
                  Significant = significant_binary[idx_ud])
    
    Z_eq = colMeans(MH_SAS$Z_post[(t/2):t, idx_eq])
    df_eq = cbind(Group = rep("Equidispersed", n_eq), 
                  Variable = rep(N, n_eq), 
                  Value = Z_eq, 
                  Significant = significant_binary[idx_eq])
    
    Z_od = colMeans(MH_SAS$Z_post[(t/2):t, idx_od])
    df_od = cbind(Group = rep("Overdispersed", n_disp), 
                  Variable = rep(N, n_disp), 
                  Value = Z_od, 
                  Significant = significant_binary[idx_od])
    
    df_long = rbind(df_long, df_ud, df_eq, df_od)
  }
}


df_summary <- df_long %>%
  group_by(Variable, Group) %>%
  summarise(mean_significance = mean(as.numeric(Significant)), .groups = "drop")

df_wide <- df_summary %>%
  pivot_wider(names_from = Variable, values_from = mean_significance)

df_long = df_long %>%
  mutate(
    Group = as.factor(Group),
    Variable = as.numeric(Variable),
    Value = as.numeric(Value),
    Significant = factor(Significant, levels = c(0,1), labels = c("Not Significant", "Significant"))
  )

```

```{r}
p = ggplot(df_long, aes(x = Group, y = Value, fill = Group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +  # suppress outliers if using jitter
  facet_wrap(~ Variable, nrow = 2) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    strip.text = element_text(face = "bold")
  ) +
  scale_fill_manual(
    values = c("Overdispersed" = "#1b9e77",  # greenish
               "Underdispersed" = "#d95f02",  # orange
               "Equidispersed" = "#1f77b4"), # purple
    name = "Group"
  ) +
  labs(
    title = "",
    y = "P(Z=1|data)",
    x = ""
  )

p

```

```{r}
ggplot(df_long, aes(x = Group, y = Value, color = Group, shape = Significant)) +
  geom_jitter(width = 0.2, height = 0, size = 1.5, alpha = 0.7) +
  facet_wrap(~ Variable, nrow = 2) +
  theme_minimal() +
  theme(
    # axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom"
  ) +
  scale_color_manual(
    values = c("Overdispersed" = "#1b9e77", 
               "Underdispersed" = "#d95f02", 
               "Equidispersed" = "#1f77b4"),
    name = "Group"
  ) +
  scale_shape_manual(
    values = c("Not Significant" = 16, "Significant" = 4),  # 16 = solid circle, 4 = x
    name = "Shape"
  ) +
  labs(
    title = "Jittered Scatterplots of simulations by varying N = number of teams in the simulated league ",
    y = "P(Z_i = 1| data)",
    x = ""
  )
```

#### Evaluation metric plots

```{r}
seasons_strvec = generate_season_string(2015, 2025)
league_acro = 'LL'
threshold = 0.5
# df_summary = data.frame()
disp_vec = list()
for (season in seasons_strvec){
  if (league_acro == 'LC' && season == '1920'){next}
  load(file = paste0("Data//MH_Results//SAS_FullLeague//",league_acro,"_",season,"_Run1_SAS.RData"))
  
  df_z1 = retrieve_nu_sas_summ(MH_SAS$Z_post, MH_SAS$nu_post, 10001, 50000)
  filter_idx = which(df_z1[1,] > threshold)
  team_filter = MH_SAS$team_names[filter_idx]
  
  disp_vec = c(disp_vec, team_filter)
}
length(disp_vec)/10
rownames(df_summary) = NULL
```

```{r}
library(patchwork)
seasons_strvec = generate_season_string(2020, 2025)
league_acro = 'LC'
threshold = 0.49
df_summary = data.frame()
for (season in seasons_strvec){
  if (league_acro == 'LC' && season == '1920'){next}
  load(file = paste0("Data//MH_Results//SAS_FullLeague//",league_acro,"_",season,"_Run1_SAS.RData"))
  
  df_z1 = retrieve_nu_sas_summ(MH_SAS$Z_post, MH_SAS$nu_post, 10001, 50000)
  filter_idx = which(df_z1[1,] > threshold)
  team_filter = MH_SAS$team_names[filter_idx]
  
  df_nu = retrieve_pars_summ(MH_SAS, c('nu'), burn_in = 0.2, c(0.1, 0.25, 0.5, 0.75, 0.9))[filter_idx, 2:6]
      
  df_tobind = cbind(rep(season, length(team_filter)),
                    team_filter,
                    df_nu)
  colnames(df_tobind) = c('season', 'team', 'q10', 'q25', 'q50', 'q75', 'q90')
  
  df_summary = rbind(df_summary, df_tobind)
}
rownames(df_summary) = NULL


# --- Assuming your df_summary looks like this:
# df_summary: season | team | q10 | q25 | q50 | q75 | q90

# --- Step 1: Take logs of numeric columns
df_summary <- df_summary %>%
  mutate(across(c(q10, q25, q50, q75, q90), log))

df_summary <- df_summary %>%
  mutate(
    color_flag = case_when(
      q25 < 0 ~ "below",
      q75 > 0 ~ "above",
      TRUE ~ "neutral"
    )
  )
# List to store plots
plot_list <- list()

# Loop through seasons
for (s in unique(df_summary$season)) {
  
  # Filter and sort teams in this season
  df_season <- df_summary %>%
    filter(season == s) %>%
    arrange(q50, q75, q25) %>%
    mutate(team = factor(team, levels = team))  # Factor levels only for this season
  
 p <- ggplot(df_season, aes(x = team, fill = color_flag)) +
    geom_boxplot(stat = "identity",
                 aes(lower = q25, middle = q50, upper = q75, ymin = q10, ymax = q90)) +
    scale_fill_manual(values = c("below" = "#1b9e77", "above" = "#d95f02")) +
    labs(title = paste0(substr(s, 1, 2), "/", substr(s, 3, 4)), x = "", y = "") +
    coord_cartesian(ylim = c(-1.5, 1.5)) +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      plot.title = element_text(face = "bold", hjust = 0.5),
      panel.background = element_rect(fill = "#f9f9f9", color = NA),
      panel.grid.major = element_line(color = "grey90"),
      panel.grid.minor = element_blank()
    )
  
  plot_list[[as.character(s)]] <- p
}

# Combine plots in 2 rows x 5 columns
combined_plot = wrap_plots(plot_list, nrow = 1, ncol = 5)
```

```{r}
seasons_strvec = generate_season_string(2020, 2025)
league_acro = 'PL'
threshold = 0.39
df_summary = data.frame()
for (season in seasons_strvec){
  if (league_acro == 'LC' && season == '1920'){next}
  load(file = paste0("Data//MH_Results//SAS_FullLeague//",league_acro,"_",season,"_Run1_SAS.RData"))
  
  df_z1 = retrieve_nu_sas_summ(MH_SAS$Z_post, MH_SAS$nu_post, 10001, 50000)
  filter_idx = which(df_z1[1,] > threshold)
  team_filter = MH_SAS$team_names[filter_idx]
  
  df_nu = retrieve_pars_summ(MH_SAS, c('nu'), burn_in = 0.2, c(0.1, 0.25, 0.5, 0.75, 0.9))[filter_idx, 2:6]
  
      
  df_tobind = cbind(rep(season, length(team_filter)),
                    team_filter,
                    df_nu)
  colnames(df_tobind) = c('season', 'team', 'q10', 'q25', 'q50', 'q75', 'q90')
  
  df_summary = rbind(df_summary, df_tobind)
}
rownames(df_summary) = NULL


# --- Assuming your df_summary looks like this:
# df_summary: season | team | q10 | q25 | q50 | q75 | q90

# --- Step 1: Take logs of numeric columns
df_summary <- df_summary %>%
  mutate(across(c(q10, q25, q50, q75, q90), log))

df_summary <- df_summary %>%
  mutate(
    color_flag = case_when(
      q25 < 0 ~ "below",
      q75 > 0 ~ "above",
      TRUE ~ "neutral"
    )
  )
# List to store plots
plot_list <- list()

# Loop through seasons
for (s in unique(df_summary$season)) {
  
  # Filter and sort teams in this season
  df_season <- df_summary %>%
    filter(season == s) %>%
    arrange(q50, q75, q25) %>%
    mutate(team = factor(team, levels = team))  # Factor levels only for this season
  
 p <- ggplot(df_season, aes(x = team, fill = color_flag)) +
    geom_boxplot(stat = "identity",
                 aes(lower = q25, middle = q50, upper = q75, ymin = q10, ymax = q90)) +
    scale_fill_manual(values = c("below" = "#1b9e77", "above" = "#d95f02")) +
    labs(title = paste0(substr(s, 1, 2), "/", substr(s, 3, 4)), x = "", y = "") +
    coord_cartesian(ylim = c(-1.5, 1.5)) +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      plot.title = element_text(face = "bold", hjust = 0.5),
      panel.background = element_rect(fill = "#f9f9f9", color = NA),
      panel.grid.major = element_line(color = "grey90"),
      panel.grid.minor = element_blank()
    )
  
  plot_list[[as.character(s)]] <- p
}

# Combine plots in 2 rows x 5 columns
combined_plot = wrap_plots(plot_list, nrow = 1, ncol = 5)
```

```{r}
#Horizontal bars

plot_list <- list()

for (s in unique(df_summary$season)) {
  
  # Filter and sort teams in this season
  df_season <- df_summary %>%
    filter(season == s) %>%
    arrange(q50, q75, q25) %>%
    mutate(team = factor(team, levels = team))  # Factor levels for this season only
  
  p <- ggplot(df_season, aes(y = team, fill = color_flag)) +
    geom_boxplot(
      stat = "identity",
      aes(
        xmin = q25, 
        xlower = q25,
        xmiddle = q50,
        xupper = q75,
        xmax = q90
      )
    ) +
    scale_fill_manual(values = c("below" = "#1b9e77", "above" = "#d95f02")) +
    labs(title = paste(s), x = "", y = "") +
    coord_cartesian(xlim = c(-1.5, 1.5)) +   # horizontal axis is VALUE axis
    theme_minimal(base_size = 12) +
    theme(
      axis.text.y = element_text(hjust = 1),
      legend.position = "none",
      plot.title = element_text(face = "bold", hjust = 0.5),
      panel.background = element_rect(fill = "#f9f9f9", color = NA),
      panel.grid.major = element_line(color = "grey90"),
      panel.grid.minor = element_blank()
    )
  
  plot_list[[as.character(s)]] <- p
}


```

```{r}

df_diff <- df_plot[1, ] - df_plot[2, ]
df_diff_long <- tibble(
  Season = names(df_diff),
  Difference = as.numeric(df_diff[1, ])
)


p3 = ggplot(df_diff_long, aes(x = Season, y = Difference, group = 1)) +
  geom_line(color = "steelblue", size = 1.2) +
  geom_point(size = 2, color = "steelblue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  theme_minimal(base_size = 14) +
  labs(title = "Difference between RPS-Pois and RPS-SAS",
       y = "Difference (Pois - SAS)")

df_diff_pct <- tibble(
  Season = names(df_plot),
  DiffPct = 100 * (df_plot[2, ] - df_plot[1, ]) / df_plot[1, ]
)

p4 = ggplot(df_diff_pct, aes(x = Season, y = DiffPct, group = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray60") +
  geom_line(color = "steelblue", size = 1.1) +
  geom_point(size = 2, color = "steelblue") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Percentage Difference between RPS-SAS and RPS-Pois",
    subtitle = "100 × (SAS − Pois) / Pois",
    x = "Season",
    y = "Difference (%)"
  )


```

```{r}
league_acros = c('PL', 'SA', 'LL', 'LC', 'BL', 'WSL')
L = 1
league_acro = league_acros[L]
game_start = 191
seas_start = 2015
seas_end = 2025
filter_type = 1
pred_type = 'oos'

outcome_ins = compute_outcome_eval(league_acro = league_acro, seas_start,seas_end, 
                                   pred_type, game_start, filter = filter_type, threshold = 0.5)
df_plot = outcome_ins[1:2, 1:(ncol(outcome_ins)-1)]
rownames(df_plot) = c('Poisson', 'CMP')

# library(tidyverse)

df_long <- as.data.frame(df_plot) %>%
  rownames_to_column("Model") %>%
  pivot_longer(
    cols = -Model,
    names_to = "Season",
    values_to = "Score"
  )

p = ggplot(df_long, aes(x = Season, y = Score, color = Model, group = Model)) +
  geom_line(size = 1.1) +
  geom_point(size = 2) +
  theme_minimal(base_size = 14) +
  labs( x = "",
       y = "RPS-Outcome")

outcome_ins = compute_exactscores_eval(league_acro = league_acro, seas_start,seas_end,pred_type, game_start, filter = filter_type, threshold = 0.5)

df_plot = outcome_ins[1:2, 1:(ncol(outcome_ins)-1)]
rownames(df_plot) = c('Poisson', 'CMP')

df_long <- as.data.frame(df_plot) %>%
  rownames_to_column("Model") %>%
  pivot_longer(
    cols = -Model,
    names_to = "Season",
    values_to = "Score"
  )

p2 = ggplot(df_long, aes(x = Season, y = Score, color = Model, group = Model)) +
  geom_line(size = 1.1) +
  geom_point(size = 2) +
  theme_minimal(base_size = 14) +
  labs( x = "Season",
       y = "RPS-ExactScore")

combined_plot = p / p2

```

```{r}

league_acros = c('PL', 'SA', 'LL', 'LC', 'BL')
# L = 1
# league_acro = league_acros[L]
game_start = 191
seas_start = 2015
seas_end = 2025

L_outcome_df = data.frame(matrix(NA_real_, nrow = 4, 5))
colnames(L_outcome_df) = league_acros
L_es_df = data.frame(matrix(NA_real_, nrow = 4, 5))
colnames(L_es_df) = league_acros
for (L in 1:5){
  league_acro = league_acros[L]
  outcome_df = compute_outcome_eval(league_acro = league_acro, seas_start, seas_end, 'oos',
                                    game_start, filter = 1, threshold = 0.5)
  es_df = compute_exactscores_eval(league_acro = league_acro, seas_start, seas_end, 'oos', 
                                   game_start, filter = 1, threshold = 0.5)
  L_outcome_df[,L] = outcome_df[,11]
  L_es_df[,L] = es_df[,11]
}

L_outcome_df[1:2,]
L_es_df[1:2,]

rbind(L_outcome_df[1:2,], L_es_df[1:2,])

```

```{r}
df_agg_filt1 = rbind(L_outcome_df[1:2,], L_es_df[1:2,])
```
